Drought indices of water demand and supply and their impact on LULCC in Chile from ERA5-Land and MODIS

### 2024-01-25 article/paper_drought_Chile2023.qmd
actualizar

title: Drought indices of water demand and supply, soil moisture, vegetation, and their impact on LULCC in continental Chile

@@

Extreme drought conditions are an important driver of tree mortality, as shown by @Senf2020 in Europe. However, we found that forest is the type of landcover macroclass less affected by variation in drought indices, being the most resistant landcover class to drought. Supporting this is @Fathi-Taperasht2022, who asserts that Indian forests are the most drought-resistant and recover rapidly. Similarly, the work of @Wu2024 who analyzed vegetation loss and recovery in response to meteorological drought in the humid subtropical Pearl River basin in China, indicates that forests showed higher drought resistance. @Xiao2023 using Vegetation Optical Depth (VOD), kNDVI, and EVI, evaluates ecosystem resistance, showing that, compared to croplands, ecosystems with a higher proportion of forests are more resilient to severe droughts. They attribute the difference to a deeper rooting depth of trees, a higher water storage capacity, and different water use strategies between them [@Xiao2023].

>>>>

title: Drought indices of water demand and supply and their impact on LULCC in Chile from ERA5-Land and MODIS

@@

Extreme drought conditions are an important driver of tree mortality, as shown by @Senf2020 in Europe. However, we found that forest is the type of landcover macroclass less affected by variation in drought indices, being the most resistant landcover class to drought. Supporting this is @Fathi-Taperasht2022, who asserts that Indian forests are the most drought-resistant and recover rapidly. Similarly, the work of @Wu2024, who analyzed vegetation loss and recovery in response to meteorological drought in the humid subtropical Pearl River basin in China, indicates that forests showed higher drought resistance. @Xiao2023, using Vegetation Optical Depth (VOD), kNDVI, and EVI, evaluates ecosystem resistance, showing that, compared to croplands, ecosystems with a higher proportion of forests are more resilient to severe droughts. They attribute the difference to a deeper rooting depth of trees, a higher water storage capacity, and different water use strategies between forest and cropland [@Xiao2023]. Despite our insights, @Venegas2022, who studied species of sclerophyllous populations, indicates that the trees experienced a general decline in growth, and the natural dynamics of their forests may have been altered, mainly due to the cumulative effects of the unprecedented drought (i.e., hydrological drought). Nonetheless, the species of sclerophyllous forest were not the predominant surface within the forest macroclass used in this study. (REVISAR ESTO, LA ultima parte lo que pasó en el bvosque esclerofilo fue el 2020, año extremo)

### 2024-01-25 article/paper_drought_Chile2023.tex
actualizar

\usepackage{lineno}\linenumbers \usepackage{multirow} \usepackage{lscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Drought indices of water demand and supply, soil moisture, vegetation, and their impact on LULCC in continental Chile},
  pdfauthor={Francisco Zambrano},
  pdfkeywords={drought, land cover change, satellite},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Drought indices of water demand and supply, soil moisture,
vegetation, and their impact on LULCC in continental Chile}
\author[1]{Francisco Zambrano%
\corref{cor1}%
\fnref{fn1}}
 \ead{francisco.zambrano@umayor.cl}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, frame hidden, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

@@

Extreme drought conditions are an important driver of tree mortality, as
shown by \citep{Senf2020} in Europe. However, we found that forest is
the type of landcover macroclass less affected by variation in drought
indices, being the most resistant landcover class to drought. Supporting
this is \citep{Fathi-Taperasht2022}, who asserts that Indian forests are
the most drought-resistant and recover rapidly. Similarly, the work of
\citep{Wu2024} who analyzed vegetation loss and recovery in response to
meteorological drought in the humid subtropical Pearl River basin in
China, indicates that forests showed higher drought resistance.
\citep{Xiao2023} using Vegetation Optical Depth (VOD), kNDVI, and EVI,
evaluates ecosystem resistance, showing that, compared to croplands,
ecosystems with a higher proportion of forests are more resilient to
severe droughts. They attribute the difference to a deeper rooting depth
of trees, a higher water storage capacity, and different water use
strategies between them \citep{Xiao2023}.

>>>>

\usepackage{lineno}\linenumbers \usepackage{multirow} \usepackage{lscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Drought indices of water demand and supply and their impact on LULCC in Chile from ERA5-Land and MODIS},
  pdfauthor={Francisco Zambrano},
  pdfkeywords={drought, land cover change, satellite},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Drought indices of water demand and supply and their impact on
LULCC in Chile from ERA5-Land and MODIS}
\author[1]{Francisco Zambrano%
\corref{cor1}%
\fnref{fn1}}
 \ead{francisco.zambrano@umayor.cl}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden]}{\end{tcolorbox}}\fi

@@

Extreme drought conditions are an important driver of tree mortality, as
shown by \citep{Senf2020} in Europe. However, we found that forest is
the type of landcover macroclass less affected by variation in drought
indices, being the most resistant landcover class to drought. Supporting
this is \citep{Fathi-Taperasht2022}, who asserts that Indian forests are
the most drought-resistant and recover rapidly. Similarly, the work of
\citep{Wu2024}, who analyzed vegetation loss and recovery in response to
meteorological drought in the humid subtropical Pearl River basin in
China, indicates that forests showed higher drought resistance.
\citep{Xiao2023}, using Vegetation Optical Depth (VOD), kNDVI, and EVI,
evaluates ecosystem resistance, showing that, compared to croplands,
ecosystems with a higher proportion of forests are more resilient to
severe droughts. They attribute the difference to a deeper rooting depth
of trees, a higher water storage capacity, and different water use
strategies between forest and cropland \citep{Xiao2023}. Despite our
insights, \citep{Venegas2022}, who studied sclerophyllous populations,
indicates that the trees experienced a general decline in growth, and
the natural dynamics of their forests may have been altered, mainly due
to the cumulative effects of the unprecedented drought (i.e.,
hydrological drought). Nonetheless, sclerophyllous forest is not the
predominant surface within the forest macroclass used in this study.
